---
title: "Automated Testing with R"
author: "David W. Body / @david_body"
date: "June 23, 2016"
output:
  ioslides_presentation:
    css: automated-testing-slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Automated Testing with R

- Who am I?
- Why are we here?
- Ask questions any time

## Automated Testing with R

### Agenda

- Why Test?
- testthat package
- Automated Testing Examples
- Automated Testing Pitfalls
- TDD Workflow (if time permits)

## Why Test?

- Increased confidence that code is correct
- Increased confidence when making changes
- Communicate intent
- Improve code structure
- Increased confidence when upgrading R and/or other packages

## testthat package

- by Hadley Wickham
- introduced in June 2011 R Journal article
- "Testing should be fun, not a drag, so you do it all the time."

Example:

```{r echo=TRUE}
library(testthat)
library(stringr)

context("String length")

test_that("str_length is number of characters", {
  expect_equal(str_length("a"), 1)
  expect_equal(str_length("ab"), 2)
  expect_equal(str_length("abc"), 3)
})
```

## testthat expectations

| Full (old style)                    | Shortcut                |
|-------------------------------------+-------------------------|
| expect_that(x, is_true())           | expect_true(x)          |
| expect_that(x, is_false())          | expect_false(x)         |
| expect_that(x, is_a(y))             | expect_is(x, y)         |
| expect_that(x, equals(y))           | expect_equal(x, y)      |
| expect_that(x, is_equivalent_to(y)) | expect_equivalent(x, y) |
| expect_that(x, is_identical_to(y))  | expect_identical(x, y)  |
| expect_that(x, matches(y))          | expect_matches(x, y)    |


## testthat expectations, cont'd

| Full (old style)                 | Shortcut             |
|----------------------------------+----------------------|
| expect_that(x, prints_text(y))   | expect_output(x, y)  |
| expect_that(x, shows_message(y)) | expect_message(x, y) |
| expect_that(x, gives_warning(y)) | expect_warning(x, y) |
| expect_that(x, throws_error(y))  | expect_error(x, y)   |


## All testhtat expectations

```{r, echo=TRUE}
match_expect <- function(name) { startsWith(name, "expect_") }
Filter(match_expect, ls("package:testthat"))
```

## Live coding example

## More detailed example

Accuracy Measures for Binary Classification

Identifying spam email

<div class="predicted"><p>predicted</p></div>
<div class="confusion">
<div class="actual"><p>actual</p></div>
<div class="confusion-main">
<table>
<tr>
<td></td>
<td class="TF">F</td>
<td class="TF">T</td>
</tr>

<tr>
<td class="TF">F</td>
<td class="cell">264</td>
<td class="cell">14</td>
</tr>

<tr>
<td class="TF">T</td>
<td class="cell">22</td>
<td class="cell">158</td>
</tr>
</table>
</div>
</div>

Zumel & Mount, Practical Data Science with R, p. 93.


## More detailed example

Accuracy Measures for Binary Classification

Identifying spam email

<div class="predicted"><p>predicted</p></div>
<div class="confusion">
<div class="actual"><p>actual</p></div>
<div class="confusion-main">
<table>
<tr>
<td></td>
<td class="TF">F</td>
<td class="TF">T</td>
</tr>

<tr>
<td class="TF">F</td>
<td class="cell">264<div class="cell-label">TN</div></td>
<td class="cell">14<div class="cell-label">FP</div></td>
</tr>

<tr>
<td class="TF">T</td>
<td class="cell">22<div class="cell-label">FN</div></td>
<td class="cell">158<div class="cell-label">TP</div></td>
</tr>
</table>
</div>
</div>

## More detailed example

Accuracy Measures for Binary Classification

Identifying spam email

<div class="predicted"><p>predicted</p></div>
<div class="confusion">
<div class="actual"><p>actual</p></div>
<div class="confusion-main">
<table>
<tr>
<td></td>
<td class="TF">F</td>
<td class="TF">T</td>
</tr>

<tr>
<td class="TF">F</td>
<td class="cell">264<div class="cell-label">TN</div></td>
<td class="cell">14<div class="cell-label">FP</div></td>
</tr>

<tr>
<td class="TF">T</td>
<td class="cell">22<div class="cell-label">FN</div></td>
<td class="cell">158<div class="cell-label">TP</div></td>
</tr>
</table>
</div>
<div class="formulas">
<p>accuracy = (TN + TP) / (TN + FP + FN + TP)</p>
</div>
</div>

## More detailed example

Accuracy Measures for Binary Classification

Identifying spam email

<div class="predicted"><p>predicted</p></div>
<div class="confusion">
<div class="actual"><p>actual</p></div>
<div class="confusion-main">
<table>
<tr>
<td></td>
<td class="TF">F</td>
<td class="TF">T</td>
</tr>

<tr>
<td class="TF">F</td>
<td class="cell">264<div class="cell-label">TN</div></td>
<td class="cell">14<div class="cell-label">FP</div></td>
</tr>

<tr>
<td class="TF">T</td>
<td class="cell">22<div class="cell-label">FN</div></td>
<td class="cell">158<div class="cell-label">TP</div></td>
</tr>
</table>
</div>
<div class="formulas">
<p>accuracy = (TN + TP) / (TN + FP + FN + TP)</p>
<p>precision = TP / (TP + FP)</p>
<p>recall = TP / (TP + FN)</p>
<p>sensitivity = TP / (TP + FN)</p>
<p>specificity = TN / (TN + FP)</p>
<p>F1 = (2 * precision * recall) / (precision + recall)</p>
</div>
</div>


## Code for more detailed example

Main code

- accuracy.R
- test_accuracy.R

Supporting code for tests

- confusion.R
- test_confusion.R


## Some automated testing pitfalls

* Controlled reproducible environment
    + setup
    + teardown
* Don't trust a test you haven't seen fail at least once
    + Make sure tests fail for the right reasons
* Don't choose examples that allow for easy mistakes
    + multiple identical values
    + needlessly symetrical matrices
    + etc
* Non-deterministic tests can create headaches


## TDD

* Test-driven development
* Example of TDD workflow (if time permits)


## Resources

Wickham, Hadley (2011). testthat: Get Started with Testing. The R Journal 3/1 5-10. http://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf

Wickham, Hadley (2015). R Packages, Testing chapter. http://r-pkgs.had.co.nz/tests.html


## Questions?

David W. Body / Big Creek Software, LLC

Twitter: [\@david_body](https://twitter.com/david_body)

[www.bigcreek.com](https://www.bigcreek.com/)
